### 2017-11-08 outline.Rmd
outline

---
title: Outline of plyranges a grammar of genomic data manipulation
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false
bibliography: references.bib
link-citations: true

---

Overview of structure, references and layout for the plyranges package paper.

# Abstract

The Bioconductor project has created many powerful abstractions for
reasoning about genomics data, such as the _Ranges_ data structures 
for representing genomic intervals. By recognising that these data structures
follow 'tidy' data principles we have created a grammar of genomic data 
manipulation that defines verbs for performing actions on and between genomic 
interval data. This grammar simplifies performing common genomic data analysis 
tasks via method chaining, type consistency and results in creating human 
readable pipelines. We have implemented this grammar as an Bioconductor/R 
package  plyranges.

# Introduction

Putting the work into context.

## What do we mean by genomic data?

  - How does that relate to 'tidy' data?
  - Why tidy data is a useful abstraction.
  - Why _Ranges_ follow this pattern.

Key references: @Wickham2014-jc, @Lawrence2013-wg

## How can we analyse genomic data?

  - the Ranges data structures
  - other abstractions for analysing genomic data
    - the query language approaches
    - the bedtools approach
    - relational algebras

Key references: @Quinlan2010-gc, @Dale2011-js, @Kaitoua2017-pw, @Ovaska2013-gd, @Mungall2014-dg @Kozanitis2016-bm

## What does plyranges contribute?

`plyranges` API provides

- a consistent grammar for analysing genomic data.
- fluent interface enables readable pipelines 
- reproducibility via code clarity, declaraitve programming,  and return consistency

Key references:

# Design and Implementation

Discussion of the API and its key features

- every action on a Range is a verb (functional composition, method chaining)
- fluent (human readable, code describes what to do rather than 'how')
- every function returns a class familiar to the user
- an expressive algebra for performing arithmetic (anchoring functions)
- split-apply-combine strategies (group_by + reduce/summarise/mutate/filter)
- recasting overlapping/nearest etc as joins

# Results

A case study showing plyranges makes life easier...
Perhaps could use long reads data from Matt and his post-doc Charity (interesting
use case of looking for intron-exon junctions)

# Conclusion (future directions)

- availability and maintenance of package
- extending to SummarisedExperiment class
- natural fit with ggbio2

# References