### 2018-05-15 README.Rmd
to generic style over plos - ready for biorxiv

---
output: github_document
bibliography: references.bib
link-citations: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tibble)
library(knitr)
library(ggbio)
library(plyranges)
library(stringr)
library(dplyr)
library(readr)
library(Biostrings)
library(AnnotationHub)
```
# plyranges-paper
The setup chunk in the [paper](paper.Rmd) shows how to download the data with R.

Requires the following packages (and versions) to reproduce the paper:

```{r, warning=FALSE}
devtools::session_info()
```