### 2018-09-20 paper.tex
new pdf

% modified from https://github.com/svmiller/svm-r-markdown-templates
\documentclass[]{article}
\usepackage[top=0.85in,right=2in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{mathpazo}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

\renewcommand*{\thefootnote}{\fnsymbol{footnote}}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

\usepackage{graphicx,grffile}
\usepackage[labelfont=bf]{caption}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}

\title{plyranges: A grammar of genomic data transformation  }

\author{\Large Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}\vspace{0.05in} \newline\normalsize\emph{}   \and \Large Dianne Cook \(^{\ref{monash}}\)\vspace{0.05in} \newline\normalsize\emph{}   \and \Large Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author}\vspace{0.05in} \newline\normalsize\emph{}  }

\date{}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\itshape}
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

% % 
% % \usepackage{biblatex}
% 
\usepackage[backend=biber, style=numeric-comp, url=false, sorting=none]{biblatex}
\addbibresource{references.bib}

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia} () and Dianne Cook \(^{\ref{monash}}\) () and Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author} ()},
             pdfkeywords = {Bioconductor, Grammar, Genomes, Data Analysis},  
            pdftitle={plyranges: A grammar of genomic data transformation},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1 
%
% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{\fontsize{18}{20}\selectfont\raggedright 
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont Stuart Lee
\footnote{\label{monash}Department of Econometrics and Business Statistics, Monash University, Clayton, Australia}\footnote{Molecular Medicine Division, Walter and Eliza Hall Institute, Parkville, Australia}} \hskip 15pt \emph{\small }   \par \textbf{\authorfont Dianne Cook \(^{\ref{monash}}\)} \hskip 15pt \emph{\small }   \par \textbf{\authorfont Michael Lawrence
\footnote{Bioinformatics and Computational Biology, Genentech Research and Early Development, South San Francisco, United States of America}\footnote{corresponding author}} \hskip 15pt \emph{\small }

}

}

\begin{abstract}

    \hbox{\vrule height .2pt width 39.14pc}

    \vskip 8.5pt % \small

\noindent The Bioconductor project provides many interoperable data abstractions
for analyzing high-throughput genomics experiments; however implementing
a typical genomic workflow with Bioconductor requires learning these
abstractions and understanding them at an integrative level. This places
a large cognitive burden on the user, especially for non-programmers. To
reduce this burden we have created a grammar of genomic data
transformation that operates on a single, central Bioconductor data
structure, GRanges, which naturally represents genomic intervals and
their associated measurements. The grammar defines verbs for performing
actions on and between genomic interval data through a simplified,
coherent interface to existing Bioconductor infrastructure, resulting in
fluent analysis workflows. We have implemented this grammar as an
R/Bioconductor package called \texttt{plyranges}.

\vskip 8.5pt \noindent \emph{Keywords}: Bioconductor, Grammar, Genomes, Data Analysis \par

    \hbox{\vrule height .2pt width 39.14pc}

\end{abstract}

\vskip 6.5pt

\noindent  \hypertarget{background}{%
\section{Background}\label{background}}

High-throughput genomics promises to unlock new disease therapies, and
strengthen our knowledge of basic biology. To deliver on those promises,
scientists must derive a stream of knowledge from a deluge of data.
Genomic data is challenging in both scale and complexity. Innovations in
sequencing technology often outstrip our capacity to process the output.
Beyond their common association with genomic coordinates, genomic data
are heterogeneous, consisting of raw sequence read alignments, genomic
feature annotations like genes and exons, and summaries like coverage
vectors, ChIP-seq peak calls, variant calls, and per-feature read
counts. Genomic scientists need software tools to wrangle the different
types of data, process the data at scale, test hypotheses, and generate
new ones, all while focusing on the biology, not the computation. For
the tool developer, the challenge is to define ways to model and operate
on the data that align with the mental model of scientists, and to
provide an implementation that scales with their ambition.

Several domain specific languages (DSLs) enable scientists to process
and reason about heterogeneous genomics data by expressing common
operations, such as range manipulation and overlap-based joins, using
the vocabulary of genomics. Their implementations either delegate
computations to a database, or operate over collections of files in
standard formats like BED. An example of the former is the Genome Query
Language (GQL) and its distributed implementation GenAp which use a
SQL-like syntax for fast retrieval of information of unprocessed
sequencing data \autocites{Kozanitis2014-va}{Kozanitis2016-bm}.
Similarly, the Genometric Query Language (GMQL) implements a DSL for
combining genomic datasets \cite{Kaitoua2017-pw}. The command line
application BEDtools develops an extensive algebra for performing
arithmetic between two or more sets of genomic regions
\cite{Quinlan2010-gc}. All of the aforementioned DSLs are designed to be
evaluated either at the command line or embedded in scripts for batch
processing. They exist in a sparse ecosystem, mostly consisting of UNIX
and database tools that lack biological semantics and operate at the
level of files and database tables.

The Bioconductor/R packages \texttt{IRanges} and \texttt{GenomicRanges}
\autocites{r-core}{Lawrence2013-wg}{Huber2015-ei} define a DSL for
analyzing genomics data with R, an interactive data analysis environment
that encourages reproducibility and provides high-level abstractions for
manipulating, modelling and plotting data, through state of the art
methods in statistical computing. The packages define object-oriented
(OO) abstractions for representing genomic data and enable
interoperability by allowing users and developers to use these
abstractions in their own code and packages. Other genomic DSLs that are
embedded in programming languages include pybedtools and valr
\autocites{Dale2011-js}{Kent2017}, however these packages lack the
interoperability provided by the aforementioned Bioconductor packages
and are not easily extended.

The Bioconductor infrastructure models the genomic data and operations
from the perspective of the power user, one who understands and wants to
take advantage of the subtle differences in data types. This design has
enabled the development of sophisticated tools, as evidenced by the
hundreds of packages depending on the framework. Unfortunately, the
myriad of data structures have overlapping purposes and important but
obscure differences in behavior that often confuse the typical end user.

Recently, there has been a concerted, community effort to standardize R
data structures and workflows around the notion of tidy data
\cite{Wickham2014-jc}. A tidy dataset is defined as a tabular data
structure that has observations as rows and columns as variables, and
all measurements pertain to a single observational unit. The tidy data
pattern is useful because it allows us to see how the data relate to the
design of an experiment and the variables measured. The \texttt{dplyr}
package \cite{Wickham2017-dplyr} defines an application programming
interface (API) that maps notions from the general relational algebra to
verbs that act on tidy data. These verbs can be composed together on one
or more tidy datasets with the pipe operator from the \texttt{magrittr}
package \cite{R-magrittr}. Taken together these features enable a user
to write human readable analysis workflows.

We have created a genomic DSL called \texttt{plyranges} that
reformulates notions from existing genomic algebras and embeds them in R
as a genomic extension of \texttt{dplyr}. By analogy, \texttt{plyranges}
is to the genomic algebra, as \texttt{dplyr} is to the relational
algebra. The \texttt{plyranges} Bioconductor package implements the
language on top of a key subset of Bioconductor data structures and thus
fully integrates with the Bioconductor framework, gaining access to its
scalable data representations and sophisticated statistical methods.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{genomic-relational-algebra}{%
\subsection{Genomic Relational
Algebra}\label{genomic-relational-algebra}}

\hypertarget{data-model}{%
\subsubsection{Data Model}\label{data-model}}

\begin{figure}
\includegraphics[width=\textwidth]{diagrams/GRanges}
\caption{An illustration of the GRanges data model for a
sample from an RNA-seq experiment. The core components of the data model
include a seqname column (representing the chromosome), a ranges column
which consists of start and end coordinates for a genomic region, and a
strand identifier (either positive, negative, or unstranded). Metadata
are included as columns to the right of the dotted line as annotations
(gene\_id) or range level covariates (score).}
\label{fig:GRanges} 
\end{figure}

The \texttt{plyranges} DSL is built on the core Bioconductor data
structure GRanges, which is a constrained table, with fixed columns for
the chromosome, start and end coordinates, and the strand, along with an
arbitrary set of additional columns, consisting of measurements or
metadata specific to the data type or experiment (figure
\ref{fig:GRanges}). GRanges balances flexibility with formal
constraints, so that it is applicable to virtually any genomic workflow,
while also being semantically rich enough to support high-level
operations on genomic ranges. As a core data structure, GRanges enables
interoperability between \texttt{plyranges} and the rest of
Bioconductor. Adhering to a single data structure simplifies the API and
makes it easier to learn and understand, in part because operations
become endomorphic, i.e., they return the same type as their input.

GRanges follow the intuitive tidy data pattern: it is a rectangular
table corresponding to a single biological context. Each row contains a
single observation and each column is a variable describing the
observations. GRanges specializes the tidy pattern in that the
observations always pertain to some genomic feature, but it largely
remains compatible with the general relational operations defined by
\texttt{dplyr}. Thus, we define our algebra as an extension of the
\texttt{dplyr} algebra, and borrow its syntax conventions and design
principles.

\begin{table}[!htbp]
\centering
\begin{tabular}{|l|l|p{6cm}|}
  \hline
  & Verb &  Description \\ 
  \hline
   & \textbf{\emph{summarize()}} & aggregate over column(s) \\ 
   Aggregate & \emph{disjoin\_ranges()} & aggregate column(s) over the union of end coordinates \\
   &  \emph{reduce\_ranges()} & aggregate column(s) by merging overlapping and neighboring ranges \\
   \hline
   &  \textbf{\emph{mutate()}} & modifies any column \\
   & \textbf{\emph{select()}} & select columns \\
  Modify (Unary) & \textbf{\emph{arrange()}} & sort by columns \\
   & \emph{stretch()} & extend range by fixed amount \\
   &  \emph{shift\_(direction)} & shift coordinates \\
   & \emph{flank\_(direction)} & generate flanking regions \\
   & \emph{\%intersection\% } & row-wise intersection \\
   & \emph{\%union\%} & row-wise union \\
   & \emph{compute\_coverage} & coverage over all ranges \\
  Modify (Binary) &  \emph{\%setdiff\%} & row-wise set difference \\
   & \emph{between()} & row-wise gap range \\
   & \emph{span()} & row-wise spanning range \\
   \hline
    & \emph{join\_overlap\_*()} & merge by overlapping ranges \\
    & \emph{join\_nearest} & merge by nearest neighbor ranges \\
    & \emph{join\_follow} & merge by following ranges \\
    Merge & \emph{join\_precedes} & merge by preceding ranges \\
    & \emph{union\_ranges} & range-wise union \\
    & \emph{intersect\_ranges} & range-wise intersect \\
    & \emph{setdiff\_ranges} & range-wise set difference \\
    & \emph{complement\_ranges} & range-wise union \\
  \hline
   & \emph{anchor\_direction()} & fix coordinates at direction \\
  Operate & \textbf{\emph{group\_by()}} & partition by column(s)  \\ 
   & \emph{group\_by\_overlaps()} & partition by overlaps \\
   \hline
   & \textbf{\emph{filter()}} & subset rows \\
  Restrict & \emph{filter\_by\_overlaps()} & subset by overlap \\
    & \emph{filter\_by\_non\_overlaps()} & subset by no overlap \\
   \hline
\end{tabular}
\caption{Overview of the \texttt{plyranges} grammar. The core verbs are
briefly described and categorized into one of the following higher level 
catgeories: aggregate, modify, merge, operate, or restrict. A verb is given bold text if
its origin is from the \texttt{dplyr} grammar.}\label{tab:grammar}
\end{table}

\hypertarget{algebraic-operations}{%
\subsubsection{Algebraic operations}\label{algebraic-operations}}

The \texttt{plyranges} DSL defines an expressive algebra for performing
genomic operations with and between GRanges objects (see table
\ref{tab:grammar}). The grammar includes several classes of operation
that cover most use cases in genomics data analysis. There are range
arithmetic operators, such as for resizing ranges or finding their
intersection, and operators for merging, filtering and aggregating by
range-specific notions like overlap and proximity.

Arithmetic operations transform range coordinates, as defined by their
\emph{start}, \emph{end} and \emph{width}. The three dimensions are
mutually dependent and partially redundant, so direct manipulation of
them is problematic. For example, changing the \emph{width} column needs
to change either the \emph{start}, \emph{end} or both to preserve
integrity of the object. We introduce the \emph{anchor} modifier to
disambiguate these adjustments. Supported anchor points include the
start, end and midpoint, as well as the 3' and 5' ends for
strand-directed ranges. For example, if we anchor the start, then
setting the width will adjust the end while leaving the start
stationary.

The algebra also defines conveniences for relative coordinate
adjustments: \emph{shift} (unanchored adjustment to both start and end)
and \emph{stretch} (anchored adjustment of width). We can perform any
relative adjustment by some combination of those two operations. The
\emph{stretch} operation requires an anchor and assumes the midpoint by
default. Since \emph{shift} is unanchored, the user specifies a suffix
for indicating the direction: left/right or, for stranded features,
upstream/downstream. For example, \emph{shift\_right} shifts a range to
the right.

The \emph{flank} operation generates new ranges that are adjacent to
existing ones. This is useful, for example, when generating upstream
promoter regions for genes. Analogous to \emph{shift}, a suffix
indicates the side of the input range to flank.

As with other genomic grammars, we define set operations that treat
ranges as sets of integers, including \emph{intersect}, \emph{union},
\emph{difference}, and \emph{complement}. There are two sets of these:
parallel and merging. For example, the parallel intersection (\emph{x
\%intersect\% y}) finds the intersecting range between \emph{xi} and
\emph{yi} for \emph{i} in \emph{1\ldots{}n}, where \emph{n} is the
length of both \emph{x} and \emph{y}. In contrast, the merging
intersection (\emph{intersect\_ranges(x, y)}) returns a new set of
disjoint ranges representing wherever there was overlap between a range
in \emph{x} and a range in \emph{y}. Finding the parallel union will
fail when two ranges have a gap, so we introduce a \emph{span} operator
that takes the union while filling any gap. The \emph{complement}
operation is unique in that it is unary. It finds the regions not
covered by any of the ranges in a single set. Closely related is the
\emph{between} parallel operation, which finds the gap separating
\emph{xi} and \emph{yi}. The binary operations are callable from within
arithmetic, restriction and aggregation expressions.

\begin{figure}
\includegraphics[width=\textwidth]{diagrams/olap-fig}
\caption{Illustration of the three overlap join operators. Each join takes two 
GRanges objects, \textit{x} and \textit{y} as input. A 'Hits' object for the 
join is computed which consists of two components. The first component contains 
the indices of the ranges in \textit{x} that have been overlapped 
(the rectangles of \textit{x} that cross the orange lines). The second 
component consists of the indices of the ranges in \textit{y} that overlap 
the ranges in \textit{x}. In this case a range in \textit{y} overlaps the 
ranges in \textit{x} three times, so the index is repeated three times. The 
resulting 'Hits' object is used to modify \textit{x} by where it was 'hit' by
\textit{y} and merge all metadata columns from \textit{x} and \textit{y} 
based on the indices contained in the 'Hits' object. This procedure is applied 
generally in the \texttt{plyranges} DSL for both overlap and nearest neighbor
operations. The join semantics alter what is returned: \textbf{A}: for an 
\textbf{inner} join the \textit{x} ranges that are overlapped by \textit{y} 
are returned. The returned ranges also include the metadata from the \textit{y}
range that overlapped the three \textit{x} ranges. \textbf{B} An 
\textbf{intersect} join is identical to an inner join except that the 
intersection is taken between the overlapped \textit{x} ranges and the 
\textit{y} ranges. \textbf{C} For the \textbf{left} join all 
\textit{x} ranges are returned regardless of whether they are overlapped by 
\textit{y}. In this case the third range (rectangle with the asterisk next to it)
of the join would have missing values on metadata columns that came from \textit{y}.}
\label{fig:olaps-fig}
\end{figure}

To support merging, our algebra recasts finding overlaps or nearest
neighbors between two genomic regions as variants of the relational join
operator. A join acts on two GRanges objects: \emph{x} and \emph{y}. The
join operator is relational in the sense that metadata from the \emph{x}
and \emph{y} ranges are retained in the joined range. All join operators
in the \texttt{plyranges} DSL generate a set of hits based on overlap or
proximity of ranges and use those hits to merge the two datasets in
different ways. There are four supported matching algorithms:
\emph{overlap}, \emph{nearest}, \emph{precede}, and \emph{follow}
(figure \ref{fig:olaps-fig}). We can further restrict the matching by
whether the query is completely \emph{within} the subject, and adding
the \emph{directed} suffix ensures that matching ranges have the same
direction (strand).

For merging based on the hits, we have three modes: \emph{inner},
\emph{intersect} and \emph{left}. The \emph{inner} overlap join is
similar to the conventional inner join in that there is a row in the
result for every match. A major difference is that the matching is not
by identity, so we have to choose one of the ranges from each pair. We
always choose the left range. The \emph{intersect} join uses the
intersection instead of the left range. Finally, the overlap \emph{left}
join is akin to left outer join in Codd's relational algebra: it
performs an overlap inner join but also returns all \emph{x} ranges that
are not hit by the \emph{y} ranges.

\begin{figure}
\includegraphics[width=\textwidth]{diagrams/code-comparison.pdf}
\caption{Idiomatic code examples for \texttt{plyranges} (\textbf{A}) and 
\texttt{GenomicRanges} (\textbf{B}) illustrating an overlap and aggregate 
operation that returns the same result. 
In each example, we have two BED files consisting of SNPs that are
genome-wide association study (GWAS) hits and reference exons. Each code block
counts for each SNP the number of distinct exons it overlaps. The 
\texttt{plyranges}code achieves this with an overlap join followed by partitioning
and aggregation. Strand is ignored by default here. 
The \texttt{GenomicRanges} code achieves this using the 'Hits' and 'List' 
classes and their methods.}
\label{fig:code}
\end{figure}

Since the GRanges object is a tabular data structure, our grammar
includes operators to filter, sort and aggregate by columns in a
GRanges. These operations can be performed over partitions formed using
the \emph{group\_by} modifier. Together with our algebra for arithmetic
and merging, these operations conform to the semantics and syntax of the
\texttt{dplyr} grammar. Consequently, \texttt{plyranges} code is
generally more compact than the equivalent \texttt{GenomicRanges} code
(figure \ref{fig:code}).

\hypertarget{developing-workflows-with-plyranges}{%
\subsection{\texorpdfstring{Developing workflows with
\texttt{plyranges}}{Developing workflows with plyranges}}\label{developing-workflows-with-plyranges}}

Here we provide illustrative examples of using the \texttt{plyranges}
DSL to show how our grammar could be integrated into genomic data
workflows. As we construct the workflows we show the data output
intermittently to assist the reader in understanding the pipeline steps.
The workflows highlight how interoperability with existing Bioconductor
infrastructure, enables easy access to public datasets and methods for
analysis and visualization.

\hypertarget{peak-finding}{%
\subsubsection{Peak Finding}\label{peak-finding}}

In the workflow of ChIP-seq data analysis, we are interested in finding
peaks from islands of coverage over chromosome. Here we will use
\texttt{plyranges} to call peaks from islands of coverage above 8 then
plot the region surrounding the tallest peak.

Using \texttt{plyranges} and the the Bioconductor package
\texttt{AnnotationHub} \cite{R-ahub} we can download and read BigWig
files from ChIP-Seq experiments from the Human Epigenome Roadmap project
\cite{Roadmap-Epigenomics-Consortium2015-pr}. Here we analyse a BigWig
file corresponding to H3 lysine 27 trimethylation (H3K27Me3) of primary
T CD8+ memory cells from peripheral blood, focussing on coverage islands
over chromosome 10.

First, we extract the genome information from the BigWig file and filter
to get the range for chromosome 10. This range will be used as a filter
when reading the file.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(plyranges)}
\NormalTok{chr10_ranges <-}\StringTok{ }\NormalTok{bw_file }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{get_genome_info}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(seqnames }\OperatorTok{==}\StringTok{ "chr10"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Then we read the BigWig file only extracting scores if they overlap
chromosome 10. We also add the genome build information to the resulting
ranges. This book-keeping is good practice as it ensures the integrity
of any downstream operations such as finding overlaps.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{chr10_scores <-}\StringTok{ }\NormalTok{bw_file }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{read_bigwig}\NormalTok{(}\DataTypeTok{overlap_ranges =}\NormalTok{ chr10_ranges) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{set_genome_info}\NormalTok{(}\DataTypeTok{genome =} \StringTok{"hg19"}\NormalTok{)}
\NormalTok{chr10_scores}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> GRanges object with 5789841 ranges and 1 metadata column:
#>             seqnames              ranges strand |              score
#>                <Rle>           <IRanges>  <Rle> |          <numeric>
#>         [1]    chr10             1-60602      * | 0.0422799997031689
#>         [2]    chr10         60603-60781      * |  0.163240000605583
#>         [3]    chr10         60782-60816      * |  0.372139990329742
#>         [4]    chr10         60817-60995      * |  0.163240000605583
#>         [5]    chr10         60996-61625      * | 0.0422799997031689
#>         ...      ...                 ...    ... .                ...
#>   [5789837]    chr10 135524723-135524734      * |  0.144319996237755
#>   [5789838]    chr10 135524735-135524775      * |  0.250230014324188
#>   [5789839]    chr10 135524776-135524784      * |  0.427789986133575
#>   [5789840]    chr10 135524785-135524806      * |  0.730019986629486
#>   [5789841]    chr10 135524807-135524837      * |   1.03103005886078
#>   -------
#>   seqinfo: 25 sequences from hg19 genome
\end{verbatim}

We then filter for regions with a coverage score greater than 8, and
following this reduce individual runs to ranges representing the islands
of coverage. This is achieved with the \texttt{reduce\_ranges()}
function, which allows a summary to be computed over each island: in
this case we take the maximum of the scores to find the coverage peaks
over chromosome 10.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{all_peaks <-}\StringTok{ }\NormalTok{chr10_scores }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{filter}\NormalTok{(score }\OperatorTok{>}\StringTok{ }\DecValTok{8}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{reduce_ranges}\NormalTok{(}\DataTypeTok{score =} \KeywordTok{max}\NormalTok{(score))}
\NormalTok{all_peaks}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> GRanges object with 1085 ranges and 1 metadata column:
#>          seqnames              ranges strand |            score
#>             <Rle>           <IRanges>  <Rle> |        <numeric>
#>      [1]    chr10     1299144-1299370      * | 13.2264003753662
#>      [2]    chr10     1778600-1778616      * | 8.20512008666992
#>      [3]    chr10     4613068-4613078      * | 8.76027011871338
#>      [4]    chr10     4613081-4613084      * | 8.43659973144531
#>      [5]    chr10             4613086      * | 8.11507987976074
#>      ...      ...                 ...    ... .              ...
#>   [1081]    chr10 135344482-135344488      * | 9.23237991333008
#>   [1082]    chr10 135344558-135344661      * |  11.843409538269
#>   [1083]    chr10 135344663-135344665      * | 8.26965999603271
#>   [1084]    chr10 135344670-135344674      * | 8.26965999603271
#>   [1085]    chr10 135345440-135345441      * | 8.26965999603271
#>   -------
#>   seqinfo: 25 sequences from hg19 genome
\end{verbatim}

Returning to the GRanges object containing normalized coverage scores,
we filter to find the coordinates of the peak containing the maximum
coverage score. We can then find a 5000 nt region centered around the
maximum position by anchoring and modifying the width.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{chr10_max_score_region <-}\StringTok{ }\NormalTok{chr10_scores }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(score }\OperatorTok{==}\StringTok{ }\KeywordTok{max}\NormalTok{(score)) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{anchor_center}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{width =} \DecValTok{5000}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

Finally, the overlap inner join is used to restrict the chromosome 10
coverage islands, to the islands that are contained in the 5000nt region
that surrounds the max peak (figure \ref{fig:peak-viz}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{peak_region <-}\StringTok{ }\NormalTok{chr10_scores }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{join_overlap_inner}\NormalTok{(chr10_max_score_region)}
\NormalTok{peak_region}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> GRanges object with 890 ranges and 2 metadata columns:
#>         seqnames            ranges strand |            score.x
#>            <Rle>         <IRanges>  <Rle> |          <numeric>
#>     [1]    chr10 21805891-21805988      * | 0.0206599999219179
#>     [2]    chr10 21805989-21806000      * | 0.0211200006306171
#>     [3]    chr10 21806001-21806044      * |  0.022069999948144
#>     [4]    chr10 21806045-21806049      * | 0.0215900000184774
#>     [5]    chr10 21806050-21806081      * | 0.0211200006306171
#>     ...      ...               ...    ... .                ...
#>   [886]    chr10          21810878      * |   5.24951982498169
#>   [887]    chr10          21810879      * |   5.83534002304077
#>   [888]    chr10 21810880-21810884      * |   6.44267988204956
#>   [889]    chr10 21810885-21810895      * |   7.07054996490479
#>   [890]    chr10 21810896-21810911      * |   6.44267988204956
#>                  score.y
#>                <numeric>
#>     [1] 29.9573001861572
#>     [2] 29.9573001861572
#>     [3] 29.9573001861572
#>     [4] 29.9573001861572
#>     [5] 29.9573001861572
#>     ...              ...
#>   [886] 29.9573001861572
#>   [887] 29.9573001861572
#>   [888] 29.9573001861572
#>   [889] 29.9573001861572
#>   [890] 29.9573001861572
#>   -------
#>   seqinfo: 25 sequences from hg19 genome
\end{verbatim}

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/peak-viz-1}

}

\caption{The final result of the \texttt{plyranges} operations to find a 5000nt region surrounding the peak of normalised coverage scores over chromosome 10, displayed as a density plot.}\label{fig:peak-viz}
\end{figure}

\hypertarget{computing-windowed-statistics}{%
\subsubsection{Computing Windowed
Statistics}\label{computing-windowed-statistics}}

Another common operation in genomics data analysis is to compute data
summaries over genomic windows. In \texttt{plyranges} this can be
achieved via the \texttt{group\_by\_overlaps()} operator. We bin and
count and find the average GC content of reads from a H3K27Me3 ChIP-seq
experiment by the Human Epigenome Roadmap Consortium.

We can directly obtain the genome information from the header of the BAM
file: in this case the reads were aligned to the hg19 genome build and
there are no reads overlapping the mitochondrial genome.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{locations <-}\StringTok{ }\NormalTok{h1_bam_sorted }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{read_bam}\NormalTok{() }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{get_genome_info}\NormalTok{() }
\end{Highlighting}
\end{Shaded}

Next we only read in alignments that overlap the genomic locations we
are interested in and select the query sequence. Note that the reading
of the BAM file is deferred: only alignments that pass the filter are
loaded into memory. We can add another column representing the GC
proportion for each alignment using the \texttt{letterFrequency()}
function from the \texttt{Biostrings} package \cite{R-biostrings}. After
computing the GC proportion as the score column, we drop all other
columns in the GRanges object.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alignments <-}\StringTok{ }\NormalTok{h1_bam_sorted }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{read_bam}\NormalTok{() }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter_by_overlaps}\NormalTok{(locations) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{select}\NormalTok{(seq) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}
    \DataTypeTok{score =} \KeywordTok{as.numeric}\NormalTok{(}\KeywordTok{letterFrequency}\NormalTok{(seq, }\StringTok{"GC"}\NormalTok{, }\DataTypeTok{as.prob =} \OtherTok{TRUE}\NormalTok{))}
\NormalTok{  ) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{select}\NormalTok{(score)}
\NormalTok{alignments}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> GRanges object with 8275595 ranges and 1 metadata column:
#>             seqnames            ranges strand |             score
#>                <Rle>         <IRanges>  <Rle> |         <numeric>
#>         [1]    chr10       50044-50119      - | 0.276315789473684
#>         [2]    chr10       50050-50119      + |              0.25
#>         [3]    chr10       50141-50213      - | 0.447368421052632
#>         [4]    chr10       50203-50278      + | 0.263157894736842
#>         [5]    chr10       50616-50690      + | 0.276315789473684
#>         ...      ...               ...    ... .               ...
#>   [8275591]     chrY 57772745-57772805      - | 0.513157894736842
#>   [8275592]     chrY 57772751-57772800      + | 0.526315789473684
#>   [8275593]     chrY 57772767-57772820      + | 0.565789473684211
#>   [8275594]     chrY 57772812-57772845      + |              0.25
#>   [8275595]     chrY 57772858-57772912      + | 0.592105263157895
#>   -------
#>   seqinfo: 24 sequences from an unspecified genome
\end{verbatim}

Finally, we create 10000nt tiles over the genome and compute the number
of reads and average GC content over all reads that fall within each
tile using an overlap join and merging endpoints.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{bins <-}\StringTok{ }\NormalTok{locations }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{tile_ranges}\NormalTok{(}\DataTypeTok{width =}\NormalTok{ 10000L)}

\NormalTok{alignments_summary <-}\StringTok{ }\NormalTok{bins }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{join_overlap_inner}\NormalTok{(alignments) }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{disjoin_ranges}\NormalTok{(}\DataTypeTok{n =} \KeywordTok{n}\NormalTok{(), }\DataTypeTok{avg_gc =} \KeywordTok{mean}\NormalTok{(score))}
\NormalTok{alignments_summary}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> GRanges object with 286030 ranges and 2 metadata columns:
#>            seqnames            ranges strand |         n            avg_gc
#>               <Rle>         <IRanges>  <Rle> | <integer>         <numeric>
#>        [1]    chr10       49999-59997      * |        88 0.369019138755981
#>        [2]    chr10       59998-69997      * |        65 0.434210526315789
#>        [3]    chr10       69998-79996      * |        56 0.386513157894737
#>        [4]    chr10       79997-89996      * |        71  0.51297257227576
#>        [5]    chr10       89997-99996      * |        64 0.387746710526316
#>        ...      ...               ...    ... .       ...               ...
#>   [286026]     chrY 57722961-57732958      * |        36 0.468201754385965
#>   [286027]     chrY 57732959-57742957      * |        38 0.469529085872576
#>   [286028]     chrY 57742958-57752956      * |        38 0.542936288088643
#>   [286029]     chrY 57752957-57762955      * |        42 0.510651629072682
#>   [286030]     chrY 57762956-57772954      * |       504 0.526942355889723
#>   -------
#>   seqinfo: 24 sequences from an unspecified genome; no seqlengths
\end{verbatim}

\hypertarget{quality-control-metrics}{%
\subsubsection{Quality Control Metrics}\label{quality-control-metrics}}

We have created a GRanges object from genotyping performed on the H1
cell line, consisting of approximately two million single nucleotide
polymorphisms (SNPs) and short insertion/deletions (indels). The GRanges
object consists of 7 columns, relating to the alleles of a SNP or indel,
the B-allele frequency, log relative intensity of the probes, GC content
score over a probe, and the name of the probe. We can use this
information to compute the transition-transversion ratio, a quality
control metric, within each chromosome in GRanges object.

First we filter out the indels and mitochondrial variants. Then we
create a logical vector corresponding to whether there is a transition
event.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{h1_snp_array <-}\StringTok{ }\NormalTok{h1_snp_array }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{filter}\NormalTok{(}\OperatorTok{!}\NormalTok{(ref }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"I"}\NormalTok{, }\StringTok{"D"}\NormalTok{)), seqnames }\OperatorTok{!=}\StringTok{ "M"}\NormalTok{) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{transition =}\NormalTok{ (ref }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"A"}\NormalTok{, }\StringTok{"G"}\NormalTok{) }\OperatorTok{&}\StringTok{ }\NormalTok{alt }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"G"}\NormalTok{,}\StringTok{"A"}\NormalTok{))}\OperatorTok{|}
\StringTok{                      }\NormalTok{(ref }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"C"}\NormalTok{,}\StringTok{"T"}\NormalTok{) }\OperatorTok{&}\StringTok{ }\NormalTok{alt }\OperatorTok{%in%}\StringTok{ }\KeywordTok{c}\NormalTok{(}\StringTok{"T"}\NormalTok{, }\StringTok{"C"}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

We then compute the transition-transversion ratio over each chromosome
using \texttt{group\_by()} in combination with \texttt{summarize()}
(figure \ref{fig:titv-viz}).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{ti_tv_results <-}\StringTok{ }\NormalTok{h1_snp_array }\OperatorTok{%>%}\StringTok{ }
\StringTok{  }\KeywordTok{group_by}\NormalTok{(seqnames) }\OperatorTok{%>%}
\StringTok{  }\KeywordTok{summarize}\NormalTok{(}\DataTypeTok{n_snps =} \KeywordTok{n}\NormalTok{(),}
            \DataTypeTok{ti_tv =} \KeywordTok{sum}\NormalTok{(transition) }\OperatorTok{/}\StringTok{ }\KeywordTok{sum}\NormalTok{(}\OperatorTok{!}\NormalTok{transition)) }
\NormalTok{ti_tv_results}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
#> DataFrame with 24 rows and 3 columns
#>     seqnames    n_snps            ti_tv
#>        <Rle> <integer>        <numeric>
#> 1          Y      2226  1.4381161007667
#> 2          6    154246 3.32013219807305
#> 3         13     83736 3.40669403220714
#> 4         10    120035 3.49400973418195
#> 5          4    153243 3.29528828096533
#> ...      ...       ...              ...
#> 20        16     77538 3.19827819589583
#> 21        12    113208 3.47851887016378
#> 22        20     57073  3.7121036988111
#> 23        21     32349 3.50480434479877
#> 24         X     55495 3.58219800181653
\end{verbatim}

\begin{figure}

{\centering \includegraphics{paper_files/figure-latex/titv-viz-1}

}

\caption{The final result of computing quality control metrics over the SNP array data with \texttt{plyranges}, displayed as a dot plot. Chromosomes are ordered by their estimated transition-transversion ratio. A white reference line is drawn at the expected ratio for a human exome.}\label{fig:titv-viz}
\end{figure}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

The design of \texttt{plyranges} adheres to well understood principles
of language and API design: cognitive consistency, cohesion,
expressiveness and endomorphism \cite{Green1996-qg}. To varying degrees,
these principles also underlie the design of \texttt{dplyr} and the
Bioconductor infrastructure.

We have aimed for \texttt{plyranges} to have a simple and direct mapping
to the user's cognitive model, i.e., how the user thinks about the data.
This requires careful selection of the level of abstraction so that the
user can express workflows in the language of genomics. This motivates
the adoption of the tidy GRanges object as our central data structure.
The basic data.frame and \texttt{dplyr} tibble lack any notion of
genomic ranges and so could not easily support our genomic grammar, with
its specific verbs for range-oriented data manipulation. Another example
of cognitive consistency is how \texttt{plyranges} is insensitive to
direction/strand by default when, e.g., detecting overlaps.
\texttt{GenomicRanges} has the opposite behavior. We believe that
defaulting to purely spatial overlap is most intuitive to most users.

To further enable cognitive consistency, \texttt{plyranges} functions
are cohesive. A function is defined to be cohesive if it performs a
singular task without producing any side-effects. Singular tasks can
always be broken down further at lower levels of abstraction. For
example, to resize a range, the user needs to specify which position
(start, end, midpoint) should be invariant over the transformation. The
\texttt{resize()} function from the \texttt{GenomicRanges} package has a
\texttt{fix} argument that sets the anchor, so calling \texttt{resize()}
coalesces anchoring and width modification. The coupling at the function
call level is justified since the effect of setting the width depends on
the anchor. However, \texttt{plyranges} increases cohesion and decouples
the anchoring into its own function call.

Increasing cohesion simplifies the interface to each operation, makes
the meaning of arguments more intuitive, and relies on function names as
the primary means of expression, instead of a more complex mixture of
function and argument names. This results in the user being able to
conceptualize the \texttt{plyranges} DSL as a flat catalog of functions,
without having to descend further into documentation to understand a
function's arguments. A flat function catalog also enhances API
discoverability, particularly through auto-completion in integrated
developer environments (IDEs). One downside of pushing cohesion to this
extreme is that function calls become coupled, and care is necessary to
treat them as a group when modifying code.

Like \texttt{dplyr}, \texttt{plyranges} verbs are functional: they are
free of side effects and are generally endomorphic, meaning that when
the input is a GRanges object they return a GRanges object. This enables
chaining of verbs through syntax like the forward pipe operator from the
\texttt{magrittr} package. This syntax has a direct cognitive mapping to
natural language and the intuitive notion of pipelines. The low-level
object-oriented APIs of Bioconductor tend to manipulate data via
sub-replacement functions, like \texttt{start(gr)\ \textless{}-\ x}.
These ultimately produce the side effect of replacing a symbol mapping
in the current environment and thus are not amenable to so-called fluent
syntax.

Expressiveness relates to the information content in code: the
programmer should be able to clarify intent without unnecessary
verbosity. For example, our overlap-based join operations are more
concise than the multiple steps necessary to achieve the same effect in
the original \texttt{GenomicRanges} API. In other cases, the
\texttt{plyranges} API increases verbosity for the sake of clarity and
cohesion. Explicitly calling \emph{anchor()} can require more typing,
but the code is easier to comprehend. Another example is the set of
routines for importing genomic annotations, including
\emph{read\_gff()}, \emph{read\_bed()}, and \emph{read\_bam()}. Compared
to the generic \emph{import()} in \texttt{rtracklayer}, the explicit
format-based naming in \texttt{plyranges} clarifies intent and the type
of data being returned. Similarly, every \texttt{plyranges} function
that computes with strand information indicates its intentions by
including suffixes such as \emph{directed}, \emph{upstream} or
\emph{downstream} in its name, otherwise strand is ignored. The
\texttt{GenomicRanges} API does not make this distinction explicit in
its function naming, instead relying on a parameter that defaults to
strand sensitivity, an arguably confusing behavior.

The implementation of \texttt{plyranges} is built on top of Bioconductor
infrastructure, meaning most functions are constructed by composing
generic functions from core Bioconductor packages. As a result, any
Bioconductor packages that uses data structures that inherit from
GRanges will be able to use \texttt{plyranges} for free. Another
consequence of building on top of Bioconductor generics is that the
speed and memory usage of \texttt{plyranges} functions are similar to
the highly optimized methods implemented in Bioconductor for GRanges
objects.

A caveat to constructing a compatible interface with \texttt{dplyr} is
that \texttt{plyranges} makes extensive use of non-standard evaluation
in R via the \texttt{rlang} package \cite{R-rlang}. Simply, this means
that computations are evaluated in the context of the GRanges objects.
Both \texttt{dplyr} and \texttt{plyranges} are based on the
\texttt{rlang} language, because it allows for more expressive code that
is free of repeated references to the container. Implicitly referencing
the container is particularly convenient when programming interactively.
Consequently, when programming with \texttt{plyranges}, a user needs to
generally understand the \texttt{rlang} language and how to adapt their
code accordingly. Users familiar with the tidyverse should already have
such knowledge.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

We have shown how to create expressive and reproducible genomic
workflows using the \texttt{plyranges} DSL. By realising that the
GRanges data model is tidy we have highlighted how to implement a
grammar for performing genomic arithmetic, aggregation, restriction and
merging. Our examples show that \texttt{plyranges} code is succinct,
human readable and can take advantage of the interoperability provided
by the Bioconductor ecosystem and the R language.

We also note that the grammar elements and design principles we have
described are programming language agnostic and could be easily be
implemented in another language where genomic information could be
represented as a tabular data structure. We chose R because it is what
we are familiar with and because the aforementioned Bioconductor
packages have implemented the GRanges data structure.

We aim to continue developing the \texttt{plyranges} package and to
extend it for use with more complex data structures, such as the
SummarizedExperiment class, the core Bioconductor data structure for
representing experimental results (e.g., counts) from multiple sample
experiments in conjunction with feature and sample metadata. Although,
the SummarizedExperiment is not strictly tidy, it does consist of three
tidy data structures that are related by feature and sample identifiers.
Therefore, the grammar and design of the \texttt{plryanges} DSL is
naturally extensible to the SummarizedExperiment.

As the \texttt{plyranges} interface encourages tidy data practices, it
integrates well with the grammar of graphics \cite{Wickham2016-gz}. To
achieve responsive performance, interactive graphics rely on lazy data
access and computing patterns, so the deferred mechanisms within
\texttt{plyranges} should help support interactive genomics
applications.

The \texttt{plyranges} package can be obtained via the Bioconductor
project website \url{https://bioconductor.org} or accessed via Github
\url{https://github.com/sa-lee/plyranges}.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{data-availability}{%
\subsection{Data Availability}\label{data-availability}}

The BigWig file for the H3K27Me3 primary T CD8+ memory cells from
peripheral blood ChIP-seq data was downloaded from the
\texttt{AnnotationHub} package (2.12.0) under accession AH33458. The BAM
file corresponding to the H1 cell line ChIP-seq data is available at GEO
under accession GSM433167. The SNP array data for the H1 cell line data
is available at GEO under accession GSM1463263.

\hypertarget{software-versions}{%
\subsection{Software Versions}\label{software-versions}}

To produce the workflows as described in results section we used R
version 3.5 with the development version of \texttt{plyranges} (1.1.10)
and the \texttt{BioStrings} package (2.48.0) installed.

All code required to reproduce this article is available at
\url{https://github.com/sa-lee/plyranges-paper}.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

We would like to thank Dr Matthew Ritchie at the Walter and Eliza Hall
Institute and Dr Paul Harrison for their feedback on earlier drafts of
this work. We would also like to thank Lori Shepherd and Hèrve Pages for
the code review they performed. This article was written with
\texttt{knitr} \cite{R-knitr} and the figures were made with
\texttt{ggbio} \cite{R-ggbio}.

\newpage
\singlespacing 
\printbibliography[title=References]

\end{document}